### 2024-05-25 thesis/input_files/data_exploratory.tex
May 25th 2024

This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies.
Finally, in order to to better understand patterns within the data, we carry out an exploratory data analysis.

\section{GEFCom2014}
The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr.\ Hong blog \cite{hong2016probabilistic}.
The main reason is that, these datasets are considered an excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
This thesis work is focused on providing forecasts for the load and price quantities.
\\
\subsection{GEFCom2014 price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consisted of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}. 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track GEFCom2014}
    \label{fig:price_track_fig1}
\end{figure}
\subsubsection{EDA}
In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see figure \ref{fig:price_track_fig1}.
After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variables, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}. From these visualisations, we can see their goodness for predicting zonal price since there exists a clear correlation between the candidate predictors and the target variables.

@@

\subsection{GEFCom2014 load track}
In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between eachother.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the hisorical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

@@

% \subsection{GEFCom ETL}
\subsubsection{GEFCom ETL}
The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
Firstly, we drop all Nans and zeros. Notice, the provided data for the challenge has a timestamp column not coherent to any time format. Thus, we first pass the data through a method extracting time information: hour, day, month and year. This info is then combined into a datetime object, we selected the ISO format.
Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables. This independent variable is named w\_avg.
The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming convention and formatting used in the last track of the challenge. The python etl pipeline has been then automated to loop every task folder through bash scripting.

\section{Germany Switzerland Austria data}
Next, we identified various data sources for Germany, Switzerland and Austria in order to further
investigate the techniques studied in this thesis work.
Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website. It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations. Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{ENTSOE} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
\\
In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
The target dependent variable is the national load while as predictors we took
the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.

In the second experiment we compared various kernels in the context of load foreacsting.
For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}. This data consists of various metereological attributes relevant for modelling electricity demand \cite{Formayer2023}. The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF). Notice, the forecasted data is splitted in two files according to the scenarios RCP4.5 and RCP8.5. In our experiment we considered the former. 
In this setting the independent variables available are 
direct irradiation, global radiation, power from reservoir plants, power from run of the river plants (ROR), weather temperature and wind potential.

@@

For what concers the data from SECURES-Met we first identified the data interesting to us. Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.

All the manipulations involved in the ETL pipeline have been carried out by combining python and bash scripting.

\subsubsection{EDA}
Following are reported the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal patter with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load (2021)}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots we can clearly see there exist a relationship that ties wind speeds and temperatures to load.

@@

Following, we used box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in figure \ref{fig:CH_day_of_week_boxplot_2021} and figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that weekdays and workdays involve a higher load and posses a wider range.

>>>>

This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our studies.
Finally, in order to better understand patterns within the data, we carry out an exploratory data analysis.

\section{GEFCom2014}
The first class of datasets we identified were the ones from the GEFCom2014 competition. The data is freely hosted on Dr.\ Hong blog \cite{hong2016probabilistic}.
The main reason is that, these datasets are considered an excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
This thesis work is focused on providing forecasts for the load and price quantities.
\\
\subsection{GEFCom2014 price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consisted of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the $1^{st}$ of January 2011 to the $15^{th}$ of June 2013, see figure \ref{fig:price_track_fig1}.\ 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track GEFCom2014}
    \label{fig:price_track_fig1}
\end{figure}
\subsubsection{EDA}
In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see figure \ref{fig:price_track_fig1}.
After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variables, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}. From these visualisations, we can see their goodness for predicting zonal price since there exists a clear correlation between the candidate predictors and the target variables.

@@

\subsection{GEFCom2014 load track}
In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between each other.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the historical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

@@

% \subsection{GEFCom ETL}
\subsubsection{GEFCom ETL}
The provided data is raw, hence the need for a standardized extract transform and load pipeline. 
Firstly, we drop all Nans and zeros. Notice, the provided data for the challenge has a timestamp column not coherent to any time format. Thus, we first pass the data through a method extracting time information: hour, day, month and year. This info is then combined into a datetime object, we selected the ISO format.
Next, since the data does not provide the location for the target utility, the best we can do is selecting the average of weather temperatures as predictor variables. This independent variable is named w avg.
The same thing applies in the cleaning of the test data with the addition of an if clause for handling the different naming convention and formatting used in the last track of the challenge. The Python ETL pipeline has been then automated to loop every task folder through Bash scripting.

\section{Germany Switzerland Austria data}
Next, we identified various data sources for Germany, Switzerland and Austria in order to further
investigate the techniques studied in this thesis work.
Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website. It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations. Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{(ENTSOE)} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
\\
In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
The target dependent variable is the national load while as predictors we took
the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.

In the second experiment we compared various kernels in the context of load forecasting.
For this study we considered the data of \href{https://zenodo.org/records/7907883}{SECURES-Met}. This data consists of various metereological attributes relevant for modelling electricity demand \cite{Formayer2023}. The dataset is made up of historical data until end of 2020, while from 2021 onwards the data consists of forecasts made by a metereological model developed by the European Centre for Medium-Range Weather Forecasts (ECMWF). Notice, the forecasted data is splitted in two files according to the scenarios RCP4.5 and RCP8.5. In our experiments we considered the former. 
In this setting the independent variables available are 
direct irradiation, global radiation, power from reservoir plants, power from run of the river plants (ROR), weather temperature and wind potential.

@@

For what concerns the data from SECURES-Met we first identified the data interesting to us. Next, we formatted all the csv to a coherent format, gave columns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.

All the manipulations involved in the ETL pipeline have been carried out by combining Python and Bash scripting.

\subsubsection{EDA}
Following are reported the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the Swiss data only, given that the plots for the German data are pretty similar.
To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal pattern with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load (2021)}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next, we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots we can clearly see there exists a relationship that ties wind speeds and temperatures to load.

@@

Following, we used box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in figure \ref{fig:CH_day_of_week_boxplot_2021} and figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that weekdays and workdays involve a higher load and possess a wider range.

### 2024-05-25 thesis/input_files/implementation.tex
22 May 2024
May 25th 2024

This section is inteneded to explain and aid for reproducibilty studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

@@

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
In our application we specified as hyperparameters the lenght of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.

\subsection{Prophet}
The \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet} model has been applied by employing the python API provided by Meta.

@@

\subsection{Support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying the liner kernel.

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented python open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point method. This algorithm is taken from the cvxopt library, see its offical manual \cite{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, absolute laplacian, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.

>>>>

This section is intended to explain and aid for reproducibility studies. Hereafter, the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of Python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.
% - indicate computer specifics
All experiments have been carried out on a 3.2 GHz 16GB Apple M1 Pro.

@@

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
In our application we specified as hyperparameters the length of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.

\subsection{Prophet}
The \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet} model has been applied by employing the Python API provided by Meta.

@@

\subsection{Support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying the linear kernel.

@@

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented Python open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C, the kernel family and its corresponding hyperparameters.
\\
In the fit method, we set up and solve the convex optimisation problem through the interior point method. This algorithm is taken from the cvxopt library, see its official manual \citeW{vandenberghe2010cvxopt} for a reference.
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the $Q$ matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: Gaussian RBF, absolute Laplacian, Matern 3/2, Matern 5/2, Linear, Cosine, Sigmoid, Periodic, Polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.

### 2024-05-25 thesis/input_files/literaturereview.tex
22 May 2024
May 25th 2024

This section presents the results of the bibliometric analysis we performed on March 6th 2024. This survey has been carried out by using the Scopus citation database. For details on the specific queries entered in Scopus, refer to Subsection 5 of the Appendix \ref{src_code}.

@@

The next question is to compare the state of point versus probabilistic forecasting, this is visualized in Figure \ref{fig:point_vs_prob}. What can be concluded is that probabilistic is less developed than point forecasting. To our mind this is due to the complexity of probabilistic forecasts.

@@

The EF literature is dominated by statistical and computational intelligence (CI) methods as can be seen from Figure \ref{fig:cs_stat_both}, with CI methods being slightly preferred. 

@@

EF is a heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked. Figure \ref{fig:subject_area} depitcs the EF publications by subject area. What can be concluded, is that the bulk of publication come from engineering, computer science, mathematics and econometrics.

@@

implementations, and applications \cite{gm_22}.

>>>>

This section presents the results of the bibliometric analysis we performed on March 6th 2024. This survey has been carried out by using the Scopus citation database. For details on the specific queries entered in Scopus, refer to Appendix \ref{src_code}.

@@

The next question is to compare the state of point-versus probabilistic forecasting, this is visualized in Figure \ref{fig:point_vs_prob}. What can be concluded is that probabilistic-is less developed than point forecasting. To our mind this is due to the complexity of probabilistic forecasts.

@@

The EF literature is dominated by statistical-and computational intelligence (CI) methods as can be seen from Figure \ref{fig:cs_stat_both}, with CI methods being slightly preferred. 

@@

EF is a heterogenous field of research, its researchers come from a wide array of backgrounds, with electrical engineers and statisticians making up the top contributors; their different educational training may explain why the split between statistical and computational intelligence methods is so marked. Figure \ref{fig:subject_area} depicts the EF publications by subject area. What can be concluded is that the bulk of publication come from engineering, computer science, mathematics and econometrics.

@@

implementations, and applications \citeW{gm_22}.
